---
title: "omr_plot_south_delta_turbidity"
author: "Catarina Pien"
date: '2023-01-04'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(dataRetrieval)
library(sharpshootR)
library(here)
source("functions.R")
```

```{r}
wytype <- read.csv("data_raw/WYType.csv") %>%
  select(WY, Basin, Yr.type) %>%
  mutate(wy = as.numeric(WY)) %>%
  pivot_wider(names_from = "Basin", values_from = Yr.type, id_cols = "wy")
```

```{r}
Sys.setenv(TZ = 'America/Los_Angeles')
```

```{r}
startDate = "2000-01-01"
endDate = "2022-12-31"
```

# Download Wind Data
* FRK, ANH, RRI, MSD, SJR, 
```{r}
wind.dir<- CDECquery("FRK", "10", "E", start = startDate,
                              end = endDate)
wind.sp<- CDECquery("FRK", "9", "E", start = startDate,
                              end = endDate)

wind_speed <- bind_rows(wind.sp)%>%
    mutate(wind_speed = value,
           datetime = ymd_hms(datetime),
           date = date(datetime),
           month = as.numeric(month)) %>%
  select(datetime,date,month, wy = water_year, station = station_id, wind_speed) %>%
  filter(!is.na(station),
         !is.na(date))

wind_dir <- bind_rows(wind.dir)%>%
    mutate(wind_dir = value,
           datetime = ymd_hms(datetime),
           date = date(datetime),
           month = as.numeric(month)) %>%
  select(datetime,date,month, wy = water_year, station = station_id, wind_dir) %>%
  filter(!is.na(station),
         !is.na(date))

wind_df <- left_join(wind_speed, wind_dir)
saveRDS(wind_cf, "data_raw/wind_FRK_2000_2022.rds")
```

# Download Flow Data
```{r}

OSJ.flow<- f_get_NWIS_flow(siteNumbers = 11313452, startDate = startDate, endDate = endDate)

flow_OSJ <- readRDS(here("data_raw", "USGS_NWIS_11313452_flow.rds")) %>%
  rename(Datetime = date_time, 
         flow = flow_inst)%>%
  mutate(station = "OSJ",
         date = date(Datetime),
         datetime = Datetime) %>%
  select(datetime, date, station, flow) %>%
  pivot_longer(cols = flow,
               names_to = "parameter",
               values_to = "flow_OSJ") 
```


# Download Turbidity Data
```{r}
stations_turb <- c("BET", "PPT", "VCU", "FRK")
stations_turb2 <- c("OBI", "HOL")

# turb <- lapply(stations_turb,
#                  function(x){
#                  CDECquery(id = x, sensor = 27, interval = "E", start = "2000-01-01",
#                               end = "2022-12-01")})
turb_df <- bind_rows(turb)%>%
    mutate(turbidity = value,
           datetime = ymd_hms(datetime),
           date = date(datetime),
           month = as.numeric(month)) %>%
  filter(turbidity>=0) %>%
  select(datetime,date,month, wy = water_year, station = station_id, turbidity) %>%
  filter(!is.na(station),
         !is.na(date))

turb2 <- lapply(stations_turb2,
                 function(x){
                 CDECquery(id = x, sensor = 221, interval = "E", start = "2000-01-01",
                              end = "2022-12-01")})
turb_df2 <- bind_rows(turb2)%>%
    mutate(turbidity = value,
           datetime = ymd_hms(datetime),
           date = date(datetime),
           month = as.numeric(month)) %>%
  filter(turbidity>0) %>%
  select(datetime,date,month, wy = water_year, station = station_id, turbidity) %>%
  filter(!is.na(station),
         !is.na(date))

turb_combined <- rbind(turb_df, turb_df2)

saveRDS(turb_combined, "data_raw/turbidity_cdec_southdelta_2000-2022.rds")
turb_combined <- readRDS("data_raw/turbidity_cdec_southdelta_2000-2022.rds")
```

# Impose some QC (Don't need to re-run)
```{r}
OBI <- turb_df2 %>%
  filter(station == "OBI",
         date == ymd("2011-02-12"))
OBI_ca <- turb_df2_ca %>%
  filter(station == "OBI",
         date == ymd("2011-02-12"))
```

## Number of values in a day
```{r}
turb_qc <- turb_combined %>%
  mutate(hour = hour(datetime)) %>%
  group_by(station, wy, month, date, hour) %>%
  slice(1) %>%
  ungroup() %>%
  select(-datetime) %>%
  group_by(station, wy, month, date) %>%
  summarise(total = n()) %>%
  mutate(Flag_QC2 = ifelse(total < 20, "Y", "N"))  %>%
  select(-total)

turb_q1 <- turb_combined %>% left_join(turb_qc) %>%
  filter(Flag_QC2 == "N") %>%
  select(datetime,date,wy, month, station , turbidity)
```

## Check for repeating values

```{r}
repeating_vals = function(df, x){
  df$same = ifelse(df$turbidity == lag(df$turbidity, 1, default = 0), 1L, 0L)
  df = df %>%
    mutate(issame = cumsum(df$same == 0L)) %>%
    group_by(station, issame) %>%
    mutate(flag = sum(same)+1 ) %>%
    ungroup() %>%
    mutate(Flag_repeats = ifelse(flag > x, "Y", "N"))
  return(df)
}
```

```{r}
# Run function repeating values and get rid of the columns we don't need
turb_q2 <- repeating_vals(df = turb_q1, x = 18) %>%
  select(-flag, -issame, -same) %>%
  rename(Flag_QC3 = Flag_repeats) 

turb_q2_clean <- turb_q2 %>%
  filter(Flag_QC3 == "N")

# Flagged values
turb_flagged <- turb_q2_clean %>%
  filter(Flag_QC3 == "Y")
```

## Remove individual high values
```{r}
turb_clean <- turb_q2_clean %>%
  filter(turbidity<200)

saveRDS(turb_clean, "data_clean/clean_turbidity_south_delta_event.rds")

```

# Calculate Daily average - Don't need to rerun
```{r}
turb_clean <- readRDS("data_clean/clean_turbidity_south_delta_event.rds")
spring_turb <- turb_clean %>%
  filter(month < 6,
         wy<2023) %>%
  group_by(date, month, wy,  station) %>%
  summarize(meanTurb = mean(turbidity)) %>%
  ungroup() %>%
  mutate(date2 = paste0(1980, "-", month, "-", day(date)),
         date2 = lubridate::ymd(date2),
         year = year(date),
         fWY = factor(wy)) %>%
 left_join(wytype) %>%
rename(wytype = SacramentoValley) %>%
  mutate(index = 1:n()) %>%
  select(index, everything())
```

## Plot data
```{r}
ggplot(spring_turb) + geom_point()

(plot_turbidity <- ggplot(spring_turb, aes(date2, meanTurb, color = fWY)) + 
    geom_line() + 
    geom_point(size = 0.8) + 
  facet_grid(station~wytype, scales = "free_y") +
  labs(y = "Mean Daily Turbidity (FNU)", color = "Water Year") + 
  viridis::scale_color_viridis(option = "turbo", discrete = TRUE) + 
  theme_bw() +
    theme(strip.text = element_text(size = 7.5),
          axis.text = element_text(size = 9),
          axis.title.x = element_blank()))
```

# More QC - outliers
```{r}
high <- spring_turb %>%
  filter(meanTurb>120)

PPThigh <- spring_turb %>%
  filter(station == "PPT" & month == 5 & wy == 2017)

PPThigh <- turb_clean %>%
  filter(station == "PPT" & month == 5 & wy == 2017)

OBIhigh <- spring_turb %>%
  filter(station == "OBI" & month == 4 & wy == 2018)
# 3966-4004  

OBIhigh <- turb_clean %>%
  filter(station == "OBI" & month == 2 & wy == 2011)

HOLhigh <- turb_clean %>%
  filter(station == "HOL" & month == 2 & wy == 2011)


FRKhigh <- turb_clean %>%
  filter(station == "FRK" & month == 1 & wy == 2021) %>%
  filter(turbidity > 50)
```

## Remove outliers
```{r}
spring_turb_clean <- spring_turb %>%
  filter(!index %in% high$index) %>%
  filter(!is.na(wytype))

saveRDS(spring_turb_clean, "data_clean/clean_turbidity_south_delta_daily.rds")
spring_turb_clean <- readRDS("data_clean/clean_turbidity_south_delta_daily.rds")
```

## Plot - Current PLOT
```{r}
(plot_turbidity <- ggplot(spring_turb_clean, aes(date2, meanTurb, color = fWY)) + 
    geom_line() + 
    geom_point(size = 0.8) + 
  facet_grid(station~wytype, scales = "free_y") +
  labs(y = "Mean Daily Turbidity (FNU)", color = "Water Year") + 
  viridis::scale_color_viridis(option = "turbo", discrete = TRUE) + 
  theme_bw() +
    theme(strip.text = element_text(size = 9),
          axis.text = element_text(size = 9.5),
          axis.text.x = element_text(angle = 90),
          axis.title.x = element_blank()))
```

```{r}
png(filename = here("figures", "omr_turbidity_2008-2022.png"), width = 8, height = 6, units = "in", pointsize = 10, family = "sans", res = 300)
plot_turbidity
dev.off()
```

# Combine with wind data and flow data 

* No flow at FRK so used OSJ

```{r}
turb_clean <- readRDS("data_clean/clean_turbidity_south_delta_event.rds")
FRK_turb <- turb_clean %>% filter(station == "FRK")
FRK_data <- left_join(wind_df, FRK_turb)
FRK_OSJ <- left_join(FRK_data, flow_OSJ %>% select(date, datetime, flow_OSJ)) %>%
  select(-Flag_QC3) 
FRK_OSJ_long <- FRK_OSJ %>%
  pivot_longer(cols = c(wind_speed:flow_OSJ), names_to = "parameter", values_to = "value")

FRK_OSJ_daily <- FRK_OSJ_long %>%
  group_by(wy, month, date,  parameter) %>%
  summarize(dailyMax = max(value, na.rm = TRUE),
            dailyMean = mean(value, na.rm = TRUE),
            dailyMin = min(value, na.rm = TRUE)) %>%
  
  filter(dailyMean<20000 & dailyMean > -20000,
         wy < 2023) 

FRK_OSJ_wide <- FRK_OSJ_daily %>%
  pivot_wider( names_from = "parameter", values_from = c("dailyMean", "dailyMax", "dailyMin"))

saveRDS(FRK_OSJ_daily, "data_clean/FRK_OSJ_wind_turbidity_flow_data.rds")
```

### Plots
```{r}
ggplot(FRK_OSJ_daily %>% filter(parameter != "wind_dir")) + 
  geom_point(aes(date, dailyMean, color = factor(wy))) + 
   geom_line(aes(date, dailyMean, color = factor(wy))) + 
  facet_wrap(~parameter, scales = "free_y", nrow = 4) +
  scale_x_date(date_breaks = "3 months") + theme_bw() + theme(axis.text.x = element_text(angle = 90))

ggplot(FRK_OSJ_daily %>% filter(parameter != "wind_dir")) + 
  geom_point(aes(date, dailyMax, color = factor(wy))) + 
   geom_line(aes(date, dailyMax, color = factor(wy))) + 
  facet_wrap(~parameter, scales = "free_y", nrow = 4) +
  scale_x_date(date_breaks = "3 months") + theme_bw() + theme(axis.text.x = element_text(angle = 90))
```


```{r}

```



# Next Steps:


* Find some instances of turbidity peaks and check them against EDSM secchi depth? 
* Check against flow and wind?? 
* Reach out to USFWS about why they use secchi?
