---
title: "clean_sacpas_data"
author: "Cat"
date: '2022-08-05'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(kableExtra)

library(scales)
library(viridis)
library("ggExtra")

```

# Read in data

Files come from SacPas: 

* MigrationTiming = (<https://www.cbr.washington.edu/sacramento/data/query_hrt.html>)[<https://www.cbr.washington.edu/sacramento/data/query_hrt.html>]
* Salvage = ("https://www.cbr.washington.edu/sacramento/data/query_salvage_hrt.html")["https://www.cbr.washington.edu/sacramento/data/query_salvage_hrt.html]

See queries:

```{r}
kbl(table <- data.frame(URL = c("MigrationTiming", "MigrationTiming", "MigrationTiming", "Salvage"),
                        Location = c("Chipps Island Trawls", "Sacramento Beach Seines Combined", "Sacramento Trawls (@ Sherwood Harbor)", NA),
                        SpeciesRun = "Chinook, Spring, Unclipped",
                        NumberYears = c(25,25,25,25),
                        Options = c("Catch Index", "Catch Index", "Catch Index", "Historical Timing"))) %>%
  kable_styling()
```


SacPas files
```{r}
filenames <- list.files("data_raw/", pattern="sacpas_unclipped_sr_chn*", full.names=TRUE)
file <- read_csv("data_raw/sacpas_unclipped_sr_chn_index_seine.csv", col_names =TRUE)

sr_list = lapply(filenames, read_csv)
sr_list2 = lapply(sr_list, `[`, -c(1:3),)
sr_mig0 <- bind_rows(sr_list2, .id = "column_label")
```

Other files
```{r}
wytype = read_csv("data_raw/WYType.csv") %>%
  janitor::clean_names()
```

# Clean data
## timing
```{r}
sr_mig <- sr_mig0 %>%
  janitor::clean_names() %>%
  mutate(across(.cols = contains("date"), lubridate::mdy)) %>%
  mutate(column_label = case_when(column_label == 1 ~ "Chipps",
                                  column_label == 2 ~ "SacBeach",
                                  column_label == 3 ~ "SacTrawlCI",
                                  column_label == 4 ~ "Salvage")) %>%
  rename(survey = column_label,
         broodyear = brood_year,
         date_first = first_passage_date,
         date_5 = x5_percent_passage_date,
         date_10 = x10_percent_passage_date,
         date_25 = x25_percent_passage_date,
         date_50 = x50_percent_passage_date,
         date_75 = x75_percent_passage_date,
         date_90 = x90_percent_passage_date, 
         date_95 = x95_percent_passage_date,
         date_last = last_passage_date) %>%
  mutate(wy = year(date_last),
         survey = as.factor(survey)) %>%
  mutate(broodyear = ifelse(is.na(broodyear), wy-1, broodyear)) %>%
  select(survey:date_last,wy, expanded_salvage) %>%
  select(survey, wy, broodyear, everything())

str(sr_mig)
```

## wy 
```{r}
wysac = filter(wytype, basin == "SacramentoValley")
```

## 95% dataset
```{r}
(nf.per <- sr_mig %>% select(survey, wy,broodyear, date_95) %>% 
   left_join(wysac %>% select(wy, yr_type, index)) %>%
  mutate(amonth = month(date_95)) %>% 
  mutate(aday = day(date_95)) %>% 
  unite(amonth, aday, col = "monthday", sep = "-") %>%
  mutate(ayear = ifelse(month(date_95) > 9, 2020, 2021)) %>%
   mutate(adate = mdy(paste0(monthday, "-", ayear))) %>%
   mutate(rdoy  = lubridate::yday(adate) + 92,
          dowy = ifelse(rdoy > 366, rdoy - 366, rdoy)))

str(nf.per)
```

## salvage
```{r}
salvage <- sr_mig %>%
  filter(survey == "Salvage") %>%
  select(wy, broodyear, expanded_salvage, date_last) %>%
  mutate(expanded_salvage = as.numeric(expanded_salvage))
```

# Calculations

## Average 95% migration date by survey and wytype
```{r}
meanDate = nf.per %>%
  group_by(survey) %>%
  summarize(meandate = ymd(mean(adate)),
            n = n()) %>%
  mutate(ayear = year(meandate)) %>%
  ungroup()

meanDateWYType = nf.per %>%
  group_by(survey, yr_type) %>%
  summarize(meandate = ymd(mean(adate)),
            n = n()) %>%
  mutate(ayear = year(meandate)) %>%
  ungroup()

dateDiff = left_join(nf.per, meanDate) %>%
  mutate(datediff = adate-meandate,
         datediff = as.numeric(datediff)) %>%
  left_join(salvage) 
```

chipps and salvage
```{r}
chipps_salvage = sr_mig %>%
  filter(survey == "Chipps") %>%
  select(wy, broodyear, date_95) %>%
  rename(date_chipps_95 = date_95) %>%
  left_join(salvage) %>%
  rename(date_last_salvage = date_last) %>%
  mutate(diffdays = date_last_salvage-date_chipps_95,
         diffdays = as.numeric(diffdays)) %>%
  left_join(wysac)
```


Long data to compare surveys to each other
```{r}
dates_wide <- dateDiff %>%
  select(survey, adate, wy, yr_type) %>%
  unique() %>%
  pivot_wider(names_from = survey, values_from = adate, id_cols = c(wy, yr_type)) %>%
  left_join(salvage) %>%
  rename(date_last_salvage = date_last)
```


# Plot data

## All dates
Timing of 95% salvage by survey
```{r}
ggplot() +
  geom_point(data = dateDiff, aes(x=adate, y= survey, color = survey), size = 2)+
  scale_x_date(date_breaks="weeks",
               labels = date_format("%m/%d")) +
  scale_color_viridis(discrete=TRUE) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle=45))
```

Timing of 95% salvage by water year type and brood year
```{r}
ggplot(nf.per, aes(x=adate, y= broodyear, color = survey, shape = yr_type))+
  geom_point(size = 10, na.rm = TRUE)+
  scale_x_date(date_breaks = "weeks",
               labels = date_format("%m/%d")) +
  scale_color_viridis(discrete = TRUE) + 
  scale_shape_manual(values=c(16, 17, 15, 6, 7))+
  theme(axis.text.x = element_text(angle=45),panel.background = element_rect(fill = 'white'),
        panel.grid.major = element_line(color = 'black'))+
    removeGrid(y=FALSE)
```

## Mean date

Mean 95% dates by water year type
```{r}
ggplot(meanDate) + geom_point(aes(x = meandate, y = survey, color = yr_type, shape = yr_type), size = 4) +
  scale_x_date(date_breaks = "1 week") +
  scale_color_viridis(discrete = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))
```

Another view (boxplot/median)
```{r}
ggplot(nf.per) + 
  geom_boxplot(aes(x = survey, y = dowy, fill = yr_type), alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_continuous(breaks = seq(100, 250, by = 10))+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))
```

How dates deviate from mean
```{r}
ggplot(dateDiff) + 
  geom_point(aes(x = wy, y = datediff, color = yr_type, shape = survey), size = 3)+
  geom_hline(yintercept = 0) +
  scale_color_viridis(discrete = TRUE) + theme_bw()
```

```{r}
ggplot(dateDiff) + 
  geom_point(aes(x = datediff, y = expanded_salvage, color = yr_type, shape = survey), size = 3) +
  #geom_smooth(aes(x = datediff, y = expanded_salvage, color = survey)) +
  geom_vline(xintercept = 0) +
  scale_color_viridis(discrete = TRUE) +
  theme_bw() 
```

## Relationships

Deviation from mean date vs. Expanded Salvage
**need some kind of standardization for expanded salvage - escapement?**

```{r}
ggplot(dates_wide) + 
  geom_point(aes(x = Salvage, y = Chipps, color = yr_type), size = 3) +
  #geom_smooth(aes(Salvage, Chipps), formula = "lm") + 
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Chipps 95% vs Salvage 95%") +
  theme_bw() 


ggplot(dates_wide) + 
  geom_point(aes(x = date_last_salvage, y = Chipps, color = yr_type), size = 3) +
  #geom_smooth(aes(x = datediff, y = expanded_salvage, color = survey)) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Chipps 95% vs last salvage")+
  theme_bw()

```

```{r}
ggplot(chipps_salvage) + 
  geom_point(aes(x = index, diffdays, color = yr_type, shape = yr_type), size = 3)+
  scale_color_viridis(discrete = TRUE) +
  theme_bw()

chipps_salvage2 <- chipps_salvage %>%
  filter(diffdays<100) 

  ggplot(chipps_salvage2) +
  geom_point(aes(x = index, y = diffdays, color = yr_type, shape = yr_type), size = 3)+
  scale_color_viridis(discrete = TRUE) +
  theme_bw()

model <- lm(index~ diffdays, data = chipps_salvage2)
summary(model)
```

# Conclusions

* Average date for 95% at Chipps: 2021-05-08
* Last 95% date for Chipps is 2021-05-23 (always before June 30)
* Last date of salvage is generally end of June (one weird year in Sept)
* Greater diff between 95% and last salvage during wetter years 


# Next steps
* Figure out relationship of grandtab with data to see if we can use proportion of escapement to compare to 95% at Chipps
* Should we compare RT monitoring group percentages with the true percentages? 
* Should we compare the real catch data with 95% at Chipps or instances of catch (tapering off)? 
* Days from June 30 by water year type
* Compare water quality with percentages? Is wt good at predicting last salvage?

# Knowns by the RT Monitoring Group
* Number fish caught so far

# Unknowns by RT Monitoring Group
* What water year type it is
* What is the total number of fish for that year