---
title: "spring_plot_turbidity"
author: "Catarina Pien"
date: '2022-12-19'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment: Plot historical turbidity for spring outflow appendix. 

* Months: Mar, Apr, May, Jun
* Sites: 
  * Chipps Island (Delta outflow; Longfin) - DWR (MAL),
  * Vernalis (Inflow) - DWR (SJR), 
  * Deepwater Shipping Channel (Delta Smelt) - USGS (DWS; 11455142),
  * Rio Vista (Delta Smelt) - USGS (SRV; 11455420)

```{r}
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(dataRetrieval)
library(CDECRetrieve) #cdec_datasets
library(here)

wytype <- read.csv("data_raw/WYType.csv") %>%
  select(WY, Basin, Yr.type) %>%
  mutate(wy = as.numeric(WY)) %>%
  pivot_wider(names_from = "Basin", values_from = Yr.type, id_cols = "wy")
```

## Get Turbidity data

NWIS Function

63680 = turbidity, NTU
```{r}
f_get_NWIS <- function(siteNumbers='11455420', parameterCd = '63680', startDate = "2000-01-01", endDate = "2022-12-01", tz = "Etc/GMT+8"){

  # get data
  print("Downloading data...")
  data <- dataRetrieval::readNWISuv(siteNumbers, parameterCd, startDate, endDate, tz)

  # fix names

  data2 <- dataRetrieval::renameNWISColumns(data)

  print("Data downloaded!")

  # clean names
  data2 <- janitor::clean_names(data2)

  # write out
  saveRDS(data2, paste0("data_raw/USGS_NWIS_", siteNumbers, "_turbidity.rds"))

  # print message!
  print("Data saved in data_raw")
}

```

Get turbidity data
```{r}
f_get_NWIS(siteNumbers = 11455142) # DWC
f_get_NWIS(siteNumbers = 11455420) # SRV

# CDEC -----------------------------------------------------------------
stations_turb <- c("MAL", "SJR")
interval <- "E" # Event = every 15 minutes, H = Hourly, A = Annual

turb <- lapply(stations_turb, 
                 function(x){
                 cdec_query(station = x, sensor_num = 27, dur_code = "E", start_date = "2000-01-01",
                              end_date = "2022-12-01")})
turb_df <- bind_rows(turb)%>%
    mutate(turb = parameter_value,
           datetime = ymd_hms(datetime),
           date = date(datetime)) %>%
  filter(turb>=0) %>%
  select(datetime,date,station = location_id, turb) %>%
  filter(!is.na(station)) 

turb_qc <- turb_df %>%
  mutate(hour = hour(datetime)) %>%
  group_by(station, date, hour) %>%
  slice(1) %>%
  ungroup() %>%
  select(-datetime) %>%
  group_by(station, date) %>%
  summarise(total = n()) %>%
  mutate(Flag_QC2 = ifelse(total < 20, "Y", "N"))  %>%
  select(-total)

turb_clean <- turb_df %>% left_join(turb_qc) %>%
  filter(Flag_QC2 == "N") %>%
  select(datetime,date,station , turb)

high <- filter(turb_clean, turb > 100 & month(date) > 2 & month(date) < 7)

```

## Get flow for QA/QC
```{r}
f_get_NWIS_flow <- function(siteNumbers=11312676, parameterCd = c('00060'), startDate = "2000-01-01", endDate = "2022-11-01", tz = "Etc/GMT+8"){

  # get data
  print("Downloading data...")
  data <- dataRetrieval::readNWISuv(siteNumbers, parameterCd, startDate, endDate, tz)

  # fix names

  data2 <- dataRetrieval::renameNWISColumns(data)

  print("Data downloaded!")

  # clean names
  data2 <- janitor::clean_names(data2)

  # write out
  saveRDS(data2, paste0("data_raw/USGS_NWIS_", siteNumbers, "_flow.rds"))
  }
```


```{r}
f_get_NWIS_flow(siteNumbers = "11303500") 
flow_Vernalis <- readRDS(here("data_raw", "USGS_NWIS_11303500_flow.rds")) %>%
  rename(Datetime = date_time, 
         flow = flow_inst)%>%
  mutate(station = "SJR",
         date = date(Datetime),
         datetime = Datetime) %>%
  select(datetime, date, station, flow) %>%
  pivot_longer(cols = flow,
               names_to = "parameter",
               values_to = "value") 

flow_vernalis_daily <- flow_Vernalis %>%
  rename(value_all = value) %>%
  group_by(date, station) %>%
  summarize(value = mean(value_all, na.rm = TRUE)) %>%
  ungroup() %>%
  rename(flow = value)

# Is there a flow station near Mallard/Chipps?? Maybe just use delta outflow
f_get_NWIS_flow(siteNumbers = "380245121532301", parameterCd = c('72137')) 


start = "2000-01-01"
end = "2022-12-01"

outflow <- cdec_query(station = "DTO" ,sensor_num = 23, dur_code = "D", start_date = start,end_date = end)

outflow_df <- bind_rows(outflow)
outflow_clean <- outflow_df %>%
    mutate(parameter = "outflow", 
           date = ymd(datetime), 
           location_id = "MAL") %>%
  select(date, station = location_id,flow = parameter_value) %>%
  filter(!is.na(flow),
         !is.na(station)) 
```



## Read in downloaded data

### Turbidity
```{r}
DWC <- readRDS(here("data_raw", "USGS_NWIS_11455142_turbidity.rds")) %>%
  rename(Datetime = date_time, 
         turbidity = bgc_project_bgc_project_ts213_ysi_exo_turb_inst,
         turbidity2 = dws_bor_hydro_project_ts213_ysi_exo_turb_inst)%>%
  mutate(station = "DWC",
         date = date(Datetime),
         datetime = Datetime,
         turb = ifelse(is.na(turbidity), turbidity2, turbidity)) %>%
  filter(turb>=0) %>%
  select(datetime, date, station, turb)

SRV <- readRDS(here("data_raw", "USGS_NWIS_11455420_turbidity.rds")) %>%
  rename(Datetime = date_time, 
         turbidity = median_ts087_ysi_6136_turb_inst,
         turbidity2 = ts213_ysi_exo_turb_inst,
         turbidity3 = lower_4_ft_from_bed_turb_inst)%>%
  mutate(station = "SRV",
         date = date(Datetime),
         datetime = Datetime,
         turb = ifelse((is.na(turbidity) & is.na(turbidity2)), turbidity3, 
                       ifelse(is.na(turbidity), turbidity2, turbidity))) %>%
  filter(turb>=0) %>%
  select(datetime, date, station, turb) 

summary(SRV)


```

#### Combine data, calculate daily mean max and min
```{r}
turbidity <- rbind(SRV, DWC, turb_clean) %>%
  group_by(date, station) %>%
  summarize(mean = mean(turb, na.rm = TRUE),
            max = max(turb, na.rm = TRUE),
            min = min(turb, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(month = month(date),
         wy = year(date),
         julian = yday(date),
         date2 = paste0(1980, "-", month, "-", day(date)),
         date2 = lubridate::ymd(date2)) %>%
  mutate(station_name = case_when(station == "DWC" ~ "DW Shipping Channel",
                                     station == "SRV" ~ "Rio Vista",
                                     station == "MAL" ~ "Mallard Island",
                                     station == "SJR" ~ "SJR Vernalis")) %>%
  filter(month>2 & month < 7,
         !is.na(date)) %>%
  ungroup() %>%
  left_join(wytype) %>%
  mutate(wytype = ifelse(station == "SJR", SanJoaquinValley, SacramentoValley)) %>%
  filter(!is.na(wytype)) %>%
  mutate(fWY = factor(wy),
         index = 1:n()) %>%
  select(index, everything())
```

## Plot data
```{r}
# Point plot by station and water year type
ggplot(turbidity, aes(date2, mean, color = fWY)) + geom_line() + geom_point(size = 0.8) + theme_bw() + facet_grid(station~wytype, scales = "free_y")

# Boxplot
ggplot(turbidity, aes(factor(month), mean, fill = station)) + geom_boxplot() + theme_bw() + facet_grid(~wytype, scales = "free_y")

```

```{r}
turbidity %>%
  filter(wytype == "BN" & mean >75) 

outflow_clean %>%
  filter(month %in% c(5,6) & year == 2010)

flows <- rbind(outflow_clean, flow_vernalis_daily)

turbidity_flow <- left_join(turbidity, flows) %>%
  filter(station %in% c("SJR", "MAL"),
         wy < 2019)

long <- turbidity_flow %>%
  rename(meanTurb = mean) %>%
  select(date, date2, wy, station, meanTurb, flow) %>%
  pivot_longer(cols = c("meanTurb", "flow"), names_to = "parameter", values_to = "value")
```

```{r}

# These look off
ggplot(long %>% filter(wy == 2010 & station == "MAL")) + 
  geom_line(aes(date, value, color = parameter)) +
  facet_wrap(~parameter, scales = "free_y", nrow = 2) + theme_bw()
ggplot(turbidity_flow %>% filter(wy == 2010 & station == "MAL")) + 
  geom_point(aes(flow, meanTurb, color = month))

# These look off
ggplot(long %>% filter(wy == 2009 & station == "SJR")) + 
  geom_line(aes(date, value, color = parameter)) +
  geom_point(aes(date, value, color = parameter)) +
  facet_wrap(~parameter, scales = "free_y", nrow = 2) + theme_bw()
ggplot(turbidity_flow %>% filter(wy == 2009 & station == "SJR")) + 
  geom_point(aes(flow, meanTurb, color = month))

# These look fine
ggplot(long %>% filter(wy == 2018 & station == "SJR")) + 
  geom_line(aes(date, value, color = parameter)) +
  facet_wrap(~parameter, scales = "free_y", nrow = 2) + theme_bw()
ggplot(turbidity_flow %>% filter(wy == 2018 & station == "SJR")) + 
  geom_point(aes(flow, meanTurb, color = month))

# Unsure - filtering this year out
ggplot(long %>% filter(wy == 2007 & station == "SJR")) + 
  geom_line(aes(date, value, color = parameter)) +
  geom_point(aes(date, value, color = parameter)) +
  facet_wrap(~parameter, scales = "free_y", nrow = 2) + theme_bw()
ggplot(turbidity_flow %>% filter(wy == 2007 & station == "SJR")) + 
  geom_point(aes(flow, meanTurb, color = month))

turb_SJR_2007 <- rbind(SRV, DWC, turb_clean) %>%
  mutate(year = year(date)) %>%
  filter(year == 2007 & station == "SJR")

ggplot(turb_SJR_2007) + geom_point(aes(x = datetime, y = turb))
```

## Remove outliers
```{r}
turbidity_f <- turbidity %>%
  filter(!(index %in% c(738, 741, 744, 411, 413))) %>%
  filter(wy > 2009) %>%
  mutate(wyclass = case_when(wytype %in% c("AN", "BN")~"Above/Below Normal",
                wytype %in% c("C", "D")~"Critical/Dry",
                wytype == "W" ~ "Wet"))
```

## Clean plot for LTO
```{r}
(plot_turbidity <- ggplot(turbidity_f, aes(date2, mean, color = fWY)) + geom_line() + geom_point(size = 0.8) + 
  facet_grid(station_name~wyclass, scales = "free_y") +
  labs(y = "Mean Daily Turbidity (FNU)", color = "Water Year") + 
  viridis::scale_color_viridis(option = "turbo", discrete = TRUE) + 
  theme_bw() +
    theme(strip.text = element_text(size = 7.5),
          axis.text = element_text(size = 9),
          axis.title.x = element_blank()))
```

```{r}
png(filename = here("figures", "turbidity_2010-2021.png"), width = 8, height = 5, units = "in", pointsize = 10, family = "sans", res = 300)
plot_turbidity
dev.off()
```

